---
title: "College Major & Income"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies {.tabset .tabset-fade}

## R
```{r r-deps, message=FALSE}
library(tidyverse)
library(scales)
```

## Python
```{python py-deps}
import pandas as pd
import numpy as np
```

# Reading data {.tabset .tabset-fade}

## R
```{r r-reading}
recent_grads <- read_csv("recent-grads.csv")
head(recent_grads)
```


## Python
```{python py-reading}
recent_grads = pd.read_csv("recent-grads.csv") # side note: entering direct raw github url resulted in some parser error.
recent_grads.head()
```

# Data cleaning {.tabset .tabset-fade}

## R 
```{r}
# cleaning step
major_processed <- recent_grads %>%
    arrange(desc(Median)) %>%
    mutate(Major = str_to_title(Major), # change all caps to 
        Major = fct_reorder(Major, Median))
```

## Python


# What categories of majors make more money than others? {.tabset .tabset-fade}


## R


## Python


# What are the highest earning majors? {.tabset .tabset-fade}


## R


## Python



# What are the lowest earning majors? {.tabset .tabset-fade}


## R


## Python



# Most common major categories  {.tabset .tabset-fade}


## R


## Python




# Most common major {.tabset .tabset-fade}


## R


## Python


# How does gender breakdown relate to typical earnings? {.tabset .tabset-fade}


## R


## Python


<!-- # Something  -->


<!-- ## R -->
<!-- ```{r} -->
<!-- # every percentage pt a field is male, the expected salary would decrease by 23650 / 100 => ~$237 -->
<!-- major_processed %>% -->
<!--     select(Major, Total, ShareWomen, Sample_size, Median) %>% -->
<!--     # linear model; Median explained by ShareWomen -->
<!--     # weighted linear regression with extra weight param -->
<!--     # it tells lm that for e.g. Petroleum Engr. > Metallurgical Engr. in Sample_size -->
<!--     lm(Median ~ ShareWomen, data = ., weight = Sample_size) %>% -->
<!--     summary() -->
<!-- ``` -->


<!-- ## Python -->








